\documentclass[sigconf]{acmart}
\acmConference{Master-Seminar}{Aktuelle Trends in der Systemsoftware}{SoSe 2026}

\settopmatter{
  printacmref=true,
  printccs=true,
  printfolios=false
}
\renewcommand\footnotetextcopyrightpermission[1]{}
\usepackage{url}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, fit, calc, backgrounds}

\lstdefinestyle{commandstyle}{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=fullflexible,
  frame=single
}

\newcommand{\parencite}[1]{\citep{#1}}

% Inline component labels matching Figure 6 colors
\newcommand{\compbox}[2]{%
  \tikz[baseline=(X.base)]%
    \node[
      rounded corners=2pt,
      inner xsep=3pt,
      inner ysep=1.5pt,
      draw=#1!70!black,
      fill=#1!12,
      line width=0.45pt
    ] (X) {\texttt{\footnotesize #2}};%
}

\newcommand{\HugeAllocatorLabel}{\compbox{blue}{HugeAllocator}}
\newcommand{\HugeCacheLabel}{\compbox{violet}{HugeCache}}
\newcommand{\HugeFillerLabel}{\compbox{green}{HugeFiller}}
\newcommand{\HugeRegionLabel}{\compbox{orange}{HugeRegion}}
\newcommand{\HugePageAwareAllocatorLabel}{\compbox{blue}{HugePageAwareAllocator}}

\definecolor{myblue}{HTML}{4C78A8}
\definecolor{mybluefill}{HTML}{EAF2FB}
\definecolor{mygreen}{HTML}{59A14F}
\definecolor{mygreenfill}{HTML}{EDF7EC}
\definecolor{myorange}{HTML}{F28E2B}
\definecolor{myorangefill}{HTML}{FDF1E3}
\definecolor{mygrayfill}{HTML}{F4F4F4}
\definecolor{myborder}{HTML}{666666}

\tikzset{
  allocused/.style={
    draw=myborder,
    fill=mybluefill,
    rounded corners=1.5pt,
    minimum height=0.62cm,
    minimum width=0.92cm,
    font=\scriptsize,
    align=center
  },
  allocfree/.style={
    draw=myborder,
    fill=mygrayfill,
    rounded corners=1.5pt,
    minimum height=0.62cm,
    minimum width=0.92cm,
    font=\scriptsize,
    align=center
  },
  component/.style={
    draw=myblue,
    fill=mybluefill,
    rounded corners=3pt,
    line width=0.8pt,
    minimum height=0.82cm,
    minimum width=2.45cm,
    font=\scriptsize,
    align=center,
    inner sep=3pt
  },
  componentAlt/.style={
    draw=mygreen,
    fill=mygreenfill,
    rounded corners=3pt,
    line width=0.8pt,
    minimum height=0.82cm,
    minimum width=2.45cm,
    font=\scriptsize,
    align=center,
    inner sep=3pt
  },
  cache/.style={
    draw=myorange,
    fill=myorangefill,
    rounded corners=3pt,
    line width=0.8pt,
    minimum height=0.72cm,
    minimum width=1.9cm,
    font=\scriptsize,
    align=center,
    inner sep=2.5pt
  },
  arrow/.style={
    -{Latex[length=2.4mm,width=1.6mm]},
    line width=0.95pt,
    draw=myborder
  },
  labelbox/.style={
    font=\scriptsize,
    text=myborder
  }
}

\begin{document}
\raggedbottom

\title{Reproducing the Temeraire Huge-Page Aware Artifact Results on a
       Local Single Docker Setup}

\author{Aldo Costa}
\email{aldo.costa@hhu.de}
\affiliation{%
  \institution{Heinrich-Heine-Universit\"at}
  \city{D\"usseldorf}
  \state{Nordrhein-Westfalen}
  \country{Deutschland}
}

\renewcommand{\shortauthors}{Costa}

\begin{abstract}
Temeraire is a hugepage-aware extension of TCMalloc whose goal is to
improve application performance by increasing hugepage coverage without
inducing uncontrolled fragmentation. The original OSDI~2021 paper
evaluates Temeraire through individual application studies, a Google
fleet experiment, and a production rollout. This report focuses on the
part of that evaluation that can be approached with public artifacts: the
Redis~6.0.9 case study.

The report first provides the memory-management background needed to
understand why hugepage-aware allocation matters, then summarizes the
relevant aspects of Temeraire's design. The reproduction artifact uses a
Docker-based setup to compare a historical public TCMalloc revision in
two configurations, one using the legacy pageheap path and one using the
hugepage-aware path. The artifact records software versions, Transparent
Huge Page settings, benchmark parameters, allocator preload state, and
raw Redis benchmark output. The result is a deliberately narrow
reproduction: it can test whether the Redis benchmark behavior is
recoverable in a local setting, but it cannot reproduce Google's
production environment or validate the paper's fleet-level claims.
\end{abstract}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10010940.10010941.10010949.10010950</concept_id>
       <concept_desc>Software and its engineering~Memory management</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10010520.10010575.10010755</concept_id>
       <concept_desc>Computer systems organization~Reliability</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Memory management}
\ccsdesc[300]{Computer systems organization~Reliability}

\keywords{memory allocation, huge pages, TCMalloc, Temeraire, Redis,
reproducibility}

\maketitle

\section{Introduction}

Memory allocation is often treated as a local implementation concern of
C and C++ programs. In large systems, however, allocation decisions also
govern where objects land in memory, whether huge pages can be exploited
effectively, and how frequently the processor must perform costly address
translation work. Hunter et al.\ make this point explicitly in the
Temeraire paper: careful allocator design can improve ``fleet
efficiency'' by reducing costs that manifest outside the allocator itself
\cite{273711}.

Temeraire modifies the pageheap of TCMalloc so that allocations are
placed with awareness of hugepage-sized regions. The allocator is
therefore not judged solely by the time spent inside \texttt{malloc} and
\texttt{free}. Its placement decisions affect subsequent execution,
because object layout influences hugepage coverage and hugepage coverage
influences TLB behavior. In some cases a more deliberate allocation
decision is worthwhile even if the decision itself carries a small
overhead.

A complete reproduction of the Temeraire paper is not feasible in a
seminar setting. The paper's strongest results come from Google's
production environment, where diverse workloads run across
warehouse-scale infrastructure and large-scale telemetry is available to
the authors. That setting is inaccessible to external researchers. The
reproduction target here is therefore deliberately smaller: the Redis
case study from the paper. Redis~6.0.9 is run under a legacy TCMalloc
pageheap configuration and under a hugepage-aware configuration, with
both variants measured across repeated \texttt{redis-benchmark} trials.

This report makes three contributions. It explains the
memory-management background needed to understand why hugepage-aware
allocation can affect application performance. It summarizes the relevant
parts of Temeraire's design. And it documents a public reproduction
artifact with pinned source versions, a containerized setup, explicit
benchmark parameters, preserved raw output, and documented deviations
from the original experimental conditions.

\section{Background}

\subsection{Paging and Address Translation}

Modern operating systems give each process its own virtual address space,
so program instructions operate on virtual rather than physical
addresses. The processor and operating system translate virtual pages to
physical page frames through page tables. Because walking these tables is
expensive, processors cache recent translations in translation lookaside
buffers (TLBs).

With ordinary 4~KiB pages, a large working set can require a large
number of distinct translations. A 2~MiB memory region, for instance,
decomposes into 512 ordinary pages; if the active working set exceeds
the TLB's effective reach, the processor must perform additional
page-table walks that stall execution and reduce application throughput.

\subsection{Huge Pages}

Huge pages reduce this address-translation overhead by mapping larger
virtual memory regions with fewer page-table entries. On x86 systems the
standard huge page size is 2~MiB, so a single TLB entry covers the same
range that would otherwise require 512 ordinary mappings. The benefit is
not that 2~MiB of data are loaded into cache at once; it is that fewer
cached translations are needed to cover the same amount of memory.

Realizing this benefit depends on memory layout. A hugepage mapping is
useful only when the virtual range is suitably aligned and can be backed
by a contiguous physical range. Linux Transparent Huge Pages (THP) can
create such mappings automatically, but the kernel's ability to do so
depends partly on the layout produced by the application allocator. When
live objects are scattered across many hugepage-sized regions, the kernel
has little opportunity to keep those regions as intact huge pages.

\subsection{Fragmentation Trade-off}

The difficulty lies in balancing memory efficiency against TLB
efficiency. Returning unused memory in small pieces can reduce resident
memory usage, but may force the kernel to split a hugepage mapping into
ordinary pages. Retaining the hugepage intact preserves TLB coverage but
keeps physical memory resident that the application no longer actively
uses.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[x=0.72cm,y=0.72cm]

    % row labels
    \node[labelbox, anchor=west] at (-0.25,2.55) {After allocation};
    \node[labelbox, anchor=west] at (-0.25,0.95) {After various free operations};

    % top row
    \foreach \i in {0,...,7} {
      \node[allocused] at (\i,1.75) {};
    }

    % bottom row
    \node[allocused] at (0,0.15) {};
    \node[allocfree] at (1,0.15) {};
    \node[allocused] at (2,0.15) {};
    \node[allocfree] at (3,0.15) {};
    \node[allocfree] at (4,0.15) {};
    \node[allocused] at (5,0.15) {};
    \node[allocfree] at (6,0.15) {};
    \node[allocused] at (7,0.15) {};

    % outer border
    \draw[draw=myborder, line width=0.9pt, rounded corners=3pt]
      (-0.75,-0.35) rectangle (7.7,2.25);

    % bottom label
    \node[labelbox] at (3.5,-0.85) {one hugepage-sized region};

    % legend below
    \node[allocused, minimum width=0.55cm, minimum height=0.38cm] at (2.4,-1.45) {};
    \node[labelbox, anchor=west] at (2.75,-1.45) {used};

    \node[allocfree, minimum width=0.55cm, minimum height=0.38cm] at (4.2,-1.45) {};
    \node[labelbox, anchor=west] at (4.55,-1.45) {free};

  \end{tikzpicture}
  \caption{Allocation and deallocation patterns can create internal
  fragmentation. Free space exists within the hugepage-sized region, but
  live allocations prevent any contiguous range from being returned to
  the OS intact. Redrawn after Figure~1 in Hunter et al.~\cite{273711}.}
  \Description{A hugepage-sized memory region before and after free operations, showing used blocks separated by fragmented free blocks.}
  \label{fig:fragmentation}
\end{figure}

Figure~\ref{fig:fragmentation} illustrates this in simplified form. The
region still contains free memory, but the free space is interrupted by
live allocations. Returning the free pieces to the OS risks losing
hugepage coverage; retaining the whole region keeps physical memory
partly idle. A hugepage-aware allocator must navigate this trade-off
explicitly, seeking to keep hot memory on huge pages while allowing
sufficiently empty regions to be released. Temeraire approaches this by
changing the placement decisions made inside TCMalloc's pageheap.

\section{Temeraire}

\subsection{TCMalloc Context}

TCMalloc is organized as a hierarchy of allocation caches. Small
allocations are typically served by per-CPU or per-thread caches before
reaching a central cache; larger, page-granularity requests eventually
reach the pageheap, which obtains memory from the OS and releases unused
spans when possible. Temeraire modifies this backend layer rather than
replacing the allocator as a whole.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[node distance=0.55cm and 0.7cm]

    \node[componentAlt] (os) {OS pageheap\\mmap / release};

    \node[cache, above=0.5cm and 0.8cm of os] (central8) {central\\8 bytes};
    \node[cache, right=0.75cm of central8] (centralbig) {central\\larger spans};

    \node[cache, above=0.65cm of central8] (transfer8) {transfer};
    \node[cache, above=0.65cm of centralbig] (transferbig) {transfer};

    \node[cache, above=0.65cm of transfer8] (cpu0) {CPU 0\\cache};
    \node[cache, above=0.65cm of transferbig] (cpu9) {CPU 9\\cache};

    \draw[arrow] (os) -- (central8);
    \draw[arrow] (os) -- (centralbig);
    \draw[arrow] (central8) -- (transfer8);
    \draw[arrow] (centralbig) -- (transferbig);
    \draw[arrow] (transfer8) -- (cpu0);
    \draw[arrow] (transferbig) -- (cpu9);

    \node[labelbox, left=0.15cm of cpu0] {malloc/free};
    \node[labelbox, below=0.48cm of os] {large requests};

    \node[
      draw=mygreen,
      dashed,
      rounded corners=4pt,
      line width=0.9pt,
      fit=(os),
      inner sep=0.22cm,
      label={[labelbox]below:Temeraire changes this backend}
    ] {};

  \end{tikzpicture}
  \caption{Simplified TCMalloc cache hierarchy. Temeraire replaces the
  pageheap backend without disturbing the upper cache layers. Redrawn
  after Figure~3 in Hunter et al.~\cite{273711}.}
  \Description{A simplified TCMalloc hierarchy from CPU caches through transfer and central caches down to the OS pageheap.}
  \label{fig:tcmalloc}
\end{figure}

Figure~\ref{fig:tcmalloc} shows the pageheap's position in the
hierarchy. Many allocation requests are resolved in the upper cache
layers and never reach the pageheap. This is architecturally convenient
for Temeraire: because the pageheap is not on the critical path for
every small allocation, it can afford more deliberate placement logic.
The cost of that logic is expected to be recovered later, through better
hugepage coverage and fewer TLB stalls in application code.

\subsection{Component Overview}

Temeraire divides the pageheap backend into several cooperating
components, shown in Figure~\ref{fig:temeraire-components}.

\begin{figure}[t]
  \centering
  \resizebox{\columnwidth}{!}{%
  \begin{tikzpicture}[
    >=Latex,
    font=\scriptsize,
    node distance=0.8cm and 1.2cm,
    every node/.style={align=center}
  ]

    \tikzset{
      allocatorbox/.style={
        draw=blue!60!black,
        fill=blue!12,
        rounded corners=4pt,
        line width=0.9pt,
        minimum width=3.0cm,
        minimum height=0.95cm,
        font=\small\ttfamily
      },
      cachebox/.style={
        draw=violet!70!black,
        fill=violet!15,
        rounded corners=4pt,
        line width=0.9pt,
        minimum width=2.55cm,
        minimum height=0.9cm,
        font=\small\ttfamily
      },
      fillerbox/.style={
        draw=green!50!black,
        fill=green!15,
        rounded corners=4pt,
        line width=0.9pt,
        minimum width=2.8cm,
        minimum height=0.9cm,
        font=\small\ttfamily
      },
      regionbox/.style={
        draw=orange!70!black,
        fill=orange!10,
        rounded corners=4pt,
        dashed,
        line width=0.9pt,
        minimum width=2.8cm,
        minimum height=0.9cm,
        font=\small\ttfamily
      },
      solidflow/.style={
        -{Latex[length=2.5mm,width=1.7mm]},
        line width=1.0pt,
        draw=black!75
      },
      dottedflow/.style={
        -{Latex[length=2.5mm,width=1.7mm]},
        dotted,
        line width=1.0pt,
        draw=black!70
      },
      textlabel/.style={
        font=\small,
        text=black
      },
      biglabel/.style={
        font=\large,
        text=black
      }
    }

    \node[allocatorbox] (allocator) {HugeAllocator};
    \node[cachebox, below=1.1cm of allocator] (cache) {HugeCache};

    \node[fillerbox, below left=2.0cm and 1.5cm of cache] (filler) {HugeFiller};
    \node[regionbox, below right=2.0cm and 1.8cm of cache] (region) {HugeRegion};

    \node[biglabel, left=3.8cm of allocator, yshift=-0.2cm] (unbacked) {unbacked hugepages};
    \node[biglabel, left=3.5cm of cache, yshift=-1.0cm] (backed) {backed hugepages};

    \draw[solidflow] (cache) -- (allocator);
    \draw[solidflow] (filler.north east) -- (cache.south west);
    \draw[dottedflow] (region.north west) -- (allocator.south east);

    \node[textlabel, below=2.0cm of filler] (smallreq)
      {small requests\\($< 1$ MiB)};
    \node[textlabel, below=2.0cm of cache] (largereq)
      {large requests\\($\geq 1$ GiB)};
    \node[textlabel, below=2.0cm of region] (medreq)
      {medium requests\\($1$ MiB -- $1$ GiB)};

    \draw[solidflow] (smallreq.north) -- (filler.south);
    \draw[solidflow] (largereq.north) -- (cache.south);
    \draw[solidflow] (medreq.north west) -- (cache.south east);
    \draw[dottedflow] (medreq.north) --
      node[right, textlabel, xshift=0.08cm] {\emph{sometimes}} (region.south);

  \end{tikzpicture}%
  }
  \caption{Temeraire's component structure. Small requests are routed to
  \texttt{HugeFiller}, large requests to \texttt{HugeCache}, and medium
  requests usually to \texttt{HugeCache} but occasionally to
  \texttt{HugeRegion}. \texttt{HugeCache} operates on backed hugepages;
  \texttt{HugeAllocator} manages unbacked hugepages. Adapted after
  Figure~6 in Hunter et al.~\cite{273711}.}
  \Description{A diagram of Temeraire components connecting HugeAllocator, HugeCache, HugeFiller, and HugeRegion with request flows by allocation size.}
  \label{fig:temeraire-components}
\end{figure}

The \HugeAllocatorLabel{} manages hugepage-aligned virtual address ranges
and supplies them to the other components on demand. The \HugeCacheLabel{}
retains fully empty, backed hugepages for reuse, avoiding the latency of
immediately releasing and re-backing memory during short-term allocation
fluctuations.

The \HugeFillerLabel{} handles small allocations. Rather than choosing
any available hole, it applies placement policies designed to preserve
useful free ranges and increase the probability that individual hugepages
eventually become completely empty, at which point they can be safely
released.

The \HugeRegionLabel{} addresses allocations too large for the filler
but too small to round to a full hugepage without significant waste.
Without this component, repeated allocations slightly above 2~MiB would
consume nearly a full hugepage each. Regions allow such allocations to
share a larger contiguous range and cross hugepage boundaries.

The top-level allocator routes requests between these components. Small
requests go to the filler, while medium and large requests are directed to
regions or full hugepage ranges depending on size. This size-based routing
reflects that small, medium, and large allocations generate qualitatively
different fragmentation risks.

\subsection{Placement Heuristic}

The central policy is implemented in \texttt{HugeFiller}. Ordinary
best-fit allocation is insufficient because Temeraire cares not only
about the total amount of free space within a hugepage, but also about
its shape. A hugepage with one long contiguous free range is more
valuable than one with many small holes, since it can accommodate larger
future requests without fragmentation.

For a request of size $K$, Temeraire prefers hugepages whose longest
free run is just sufficient for the request, thereby avoiding the
unnecessary consumption of longer ranges. When multiple hugepages share
the same longest-free-run status, the allocator prefers the fuller one,
using allocation count as a proxy for proximity to becoming empty. The
intuition is that a hugepage with few live allocations is more likely to
drain naturally; disturbing it with a new allocation reduces that
probability.

This heuristic captures the central design idea. Placement decisions are
deliberately more conservative than a naive strategy because poor
placement creates long-lived fragmentation, whereas careful placement
preserves hugepage coverage and reduces subsequent TLB stalls. The
pageheap's position outside the hot allocation path makes this additional
reasoning affordable.

\section{Related Work}

Prior work on memory allocation and huge pages divides broadly into
allocator-side and OS-side approaches. Temeraire belongs to the former
category: it changes object placement in user space rather than relying
on the kernel to recover hugepage-friendly layouts after the fact.

SuperMalloc is a scalable multithreaded allocator for 64-bit systems
whose primary goals are allocation speed and scalability
\parencite{10.1145/2754169.2754178}. It uses huge pages for very large
allocations, but this use is narrower than Temeraire's: SuperMalloc does
not attempt to make the general pageheap layout hugepage-aware across a
broad range of allocation sizes, nor does it address the fragmentation
that arises when small and medium-sized allocations interact with THP.

MallocPool targets TLB efficiency through a different abstraction
\parencite{8502619}. It preallocates a large memory region and serves
chunks from it contiguously, allowing dynamically allocated objects to
benefit from favorable TLB mappings. While related to Temeraire's focus
on memory layout, MallocPool is not integrated into a general-purpose
pageheap and does not need to balance the same combination of allocation
sizes, release decisions, and production server workloads.

LLAMA uses learning-based lifetime prediction for C++ server workloads
\parencite{10.1145/3373376.3378525}. It co-locates allocations by
predicted lifetime on the grounds that huge pages are easier to release
when all resident objects have compatible lifetimes. Mixing short-lived
and long-lived objects on the same hugepage makes that page difficult to
reclaim. Temeraire targets a related problem without lifetime prediction,
relying instead on online placement heuristics, principally
longest-free-range information and allocation count, that are available
at allocation time without profiling.

On the OS side, Ingens and HawkEye manage huge pages from the kernel
\parencite{10.5555/3026877.3026931,10.1145/3297858.3304064}. These
systems improve hugepage allocation, distribution, or reclamation without
modifying the application allocator, making them complementary to
Temeraire rather than competing with it. Kernel-level management is
constrained by the virtual memory layout exposed from user space: if the
allocator scatters live objects across hugepage-sized regions, the kernel
faces fragmentation and memory pressure regardless of its own policies.

Taken together, these approaches illustrate why huge pages are attractive
but difficult to exploit reliably. Temeraire's contribution is to make
hugepage coverage an explicit objective inside a production allocator
backend, packing some hugepages densely while allowing others to drain to
the point where release is worthwhile.

\section{Original Evaluation Methodology}

The original paper evaluates Temeraire at several levels of granularity.
Application-level results cover throughput, latency, memory usage,
instructions per cycle, TLB-related stalls, and allocation overhead for
individual production workloads. A Redis case study provides a more
accessible comparison because Redis is open-source and its workload can
be described precisely. Beyond these application experiments, the authors
report a controlled fleet experiment and longitudinal data from a
production rollout.

These levels answer distinct questions. The Redis case study asks whether
the allocator change affects a concrete, publicly available application
under controlled benchmark conditions. The fleet experiment and rollout
ask whether hugepage-aware allocation improves aggregate efficiency
across diverse warehouse-scale workloads. Only the former is amenable to
external reproduction, and this artifact targets it exclusively.

The paper reports that the Redis experiment used Redis~6.0.9 and
compared Temeraire with TCMalloc's legacy pageheap across repeated
\texttt{redis-benchmark} trials. Each trial issued 1,000,000 requests
combining a five-element push and a five-element read. These parameters
are sufficiently explicit to form the basis of a local reproduction
attempt.

\section{Reproduction Scope}

This artifact reproduces only the Redis case study. The fleet-scale
results from the paper rely on Google's internal workload diversity,
telemetry infrastructure, and production hardware, none of which are
publicly accessible. The benchmark can therefore characterize how the
Redis workload behaves on the tested machine under two allocator
configurations; it cannot confirm the paper's aggregate efficiency
claims, nor is it intended to.

\section{Artifact Design}

The repository is organized around a Docker container that provides a
Linux environment with compiler tools, Redis tooling, \texttt{perf}, and
the required build dependencies. The host directory is mounted into the
container so that raw outputs remain available in the repository tree
after each run.

The setup script pins all relevant public source versions:

\begin{itemize}
  \item Redis \texttt{6.0.9};
  \item \texttt{gperftools-2.16}, used only as an optional side
        baseline;
  \item public \texttt{google/tcmalloc} commit \texttt{8e534f507074}.
\end{itemize}

The selected TCMalloc revision is a practical approximation. It still
exposes the \texttt{want\_no\_hpaa} mechanism needed to build a
legacy-pageheap variant, but the public repository does not provide two
ready-to-preload shared libraries for this experiment. Since the
benchmark switches allocator modes through \texttt{LD\_PRELOAD}, both
configurations must be packaged as loadable shared objects.

To that end, the setup adds a small wrapper target inside the cloned
TCMalloc tree. The wrapper contributes no allocator logic; it only
provides the linkage needed for two shared libraries. One variant links
against \path{//tcmalloc}. The legacy variant also links against
\path{//tcmalloc:want_no_hpaa}. The benchmark scripts then select
\path{libtcmalloc_temeraire.so} or \path{libtcmalloc_legacy.so} at
runtime through \path{LD_PRELOAD}.

This arrangement should not be read as a reconstruction of Google's
internal allocator artifact. It is a public-code comparison that
preserves the legacy-versus-hugepage-aware distinction at a level
sufficient for a local Redis experiment.

\section{Reproduction Methodology}

The main workflow is:

\begin{lstlisting}[style=commandstyle]
docker compose build
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/setup_env.sh"
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/check_allocator_preload.sh"
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/collect_system_info.sh"
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/run_redis_benchmark.sh legacy"
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/run_redis_benchmark.sh temeraire"
\end{lstlisting}

For the final paper comparison, the repository also provides a
paper-closer orchestration path:

\begin{lstlisting}[style=commandstyle]
docker compose run --rm temeraire-dev bash -lc \
  "echo always > /sys/kernel/mm/transparent_hugepage/enabled"
docker compose run --rm temeraire-dev bash -lc \
  "echo always > /sys/kernel/mm/transparent_hugepage/defrag"
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/run_paper_closer_redis_experiment.sh"
\end{lstlisting}

Repeated paper-closer runs can use a balanced allocator order:

\begin{lstlisting}[style=commandstyle]
docker compose run --rm temeraire-dev bash -lc \
  "./scripts/run_paper_closer_redis_experiment.sh --allocator-order balanced"
\end{lstlisting}

This alternates which allocator is measured first across repeated runs and
records the effective order in the manifest. It was added because the expected
Redis effect is below one percent, so a fixed allocator order can make host
drift harder to separate from allocator behavior.

The benchmark script starts Redis, selects the allocator mode, records
memory state, and executes repeated benchmark trials. Each trial flushes
the database and then performs two operations:

\begin{itemize}
  \item \texttt{LPUSH benchmark:list v1 v2 v3 v4 v5};
  \item \texttt{LRANGE benchmark:list 0 4}.
\end{itemize}

The default configuration follows the paper-shaped Redis workload: 2000
trials, 1,000,000 requests per trial, 50 clients, and pipeline depth~16.
The paper-closer script runs this workload in two modes, periodic
release off and periodic release on, to match the two Redis rows in
Table~1 of Hunter et al.: \textit{Redis\textsuperscript{\textdagger}}
(periodic release disabled) and \textit{Redis} (periodic release
enabled). Results from the optional smoke-test mode, which reduces trial
and request counts for environment validation, must not be compared
against the paper.

Reproducibility metadata is collected prior to each benchmark run,
covering kernel version, OS release, CPU model, memory summary, THP
settings, Redis version, TCMalloc commit, and allocator artifacts. The
allocator preload check is particularly important because Redis may
report \texttt{mem\_allocator:~libc} even when \texttt{LD\_PRELOAD} has
substituted a different allocator at the process level.

The paper-closer rerun also confirmed that the LLVM toolchain path was
exercised correctly: Redis and the allocator wrapper were compiled
against a locally built LLVM~13.0.0 from commit
\texttt{cd442157cff4aad209ae532cbf031abbe10bc1df}. This closes one
methodological gap relative to the paper, though differences in host
hardware, hypervisor, and kernel behavior remain unavoidable.

A further complication arose during the initial paper-closer attempts.
With the host in THP \texttt{madvise} mode under WSL2, process memory
summaries showed no hugepage-backed anonymous memory throughout the
entire run. The final rerun therefore set the shared WSL2/Docker THP
policy to \texttt{always} before launching the benchmark, and added
periodic \texttt{smaps\_rollup} sampling so that hugepage activity could
be observed mid-run rather than only in pre- and post-run snapshots.

The primary measured quantity is throughput in requests per second as
reported by \texttt{redis-benchmark}. Raw benchmark logs are preserved
in the artifact so that per-trial variance can be inspected directly.
Where \texttt{perf} is available on the host, it can additionally
collect hardware events including cycles, instructions, page faults, and
TLB-related counters. These must be interpreted cautiously, however,
given the effects of virtualization, container permissions, and counter
multiplexing on reported values.

\section{Results}

The final analysis uses only runs that exercised the intended THP path. An
earlier paper-closer attempt, performed while the WSL2 kernel exposed THP in
\texttt{madvise} mode, recorded \texttt{AnonHugePages: 0 kB} for Redis and is
therefore treated as workflow validation rather than as paper-comparable data.
After switching the shared WSL2/Docker THP policy to \texttt{always}, every
reported run showed non-zero \texttt{AnonHugePages} in Redis process snapshots.

For each Redis run, the artifact records separate LPUSH and LRANGE throughput.
The comparison below uses a combined push-plus-read metric. For each trial it
computes
\[
  2 / (1/r_{\text{LPUSH}} + 1/r_{\text{LRANGE}})
\]
and then averages across trials. This harmonic-mean formulation avoids treating
the two operations as independent paper results, since Hunter et al.\ report
only one Redis throughput delta per release mode.

\begin{table}[ht]
  \caption{Run-level Temeraire throughput deltas in THP-enabled runs. L first
           and T first denote whether legacy or Temeraire ran first within the
           allocator pair. off/on denote periodic release disabled/enabled.}
  \label{tab:run-deltas}
  \footnotesize
  \begin{tabular}{@{}lccc@{}}
    \toprule
    Run & Order & off & on \\
    \midrule
    fixed-order & L first & $+1.88\%$ & $+0.27\%$ \\
    balanced 1 & L first & $+0.44\%$ & $+0.12\%$ \\
    balanced 2 & T first & $+0.48\%$ & $+0.38\%$ \\
    balanced 3 & L first & $-0.76\%$ & $-2.37\%$ \\
    balanced 4 & T first & $+3.51\%$ & $-12.07\%$ \\
    targeted on & T first & -- & $+0.12\%$ \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:run-deltas} shows that release-off has the clearer local signal:
four of five THP-enabled full runs favor Temeraire, and the median release-off
delta is $+0.48\%$. This is below, but close to, the paper's
$+0.75\%$ Redis\textsuperscript{\textdagger} result. The release-on case is
weaker. Most release-on runs are near flat or modestly positive, but one
balanced run produced a large negative result.

That large negative release-on result was investigated separately. The
Temeraire run from balanced rerun~4 had correct metadata, background release
enabled, active THP, stable RSS, stable \texttt{AnonHugePages}, and a clean
Redis log. Its slowdown was visible from the first 250-trial window and only
partially recovered at the end of the run, while the paired legacy run was
stable. A targeted release-on-only rerun with Temeraire first then returned to
$+0.12\%$ combined throughput: LPUSH was $-0.04\%$ and LRANGE was $+0.37\%$.
This supports treating the $-12.07\%$ release-on result as a transient
single-host outlier rather than as representative allocator behavior.

\begin{table}[ht]
  \caption{Paper comparison and local interpretation.}
  \label{tab:paper-compare}
  \footnotesize
  \begin{tabular}{@{}lccp{0.34\columnwidth}@{}}
    \toprule
    Mode & Paper & Local summary & Interpretation \\
    \midrule
    off & $+0.75\%$ & median $+0.48\%$ & recurring positive signal \\
    on  & $+0.44\%$ & about $0$ to $+0.38\%$ & small and noisy \\
    \bottomrule
  \end{tabular}
\end{table}

The result is therefore mixed but informative. The artifact does not reproduce
the exact Redis Table~1 statistics. It does, however, repeatedly put the local
Redis comparison in the paper's small-effect regime once THP is actually
active. Release-off is the stronger reproduction signal. Release-on remains
close to the paper's reported scale in most runs, but the single-host setup is
noisy enough that the effect should be described as near-flat rather than
robustly reproduced.

\section{Discussion}

The strongest aspect of this artifact is procedural reproducibility. It
records exact public source versions, benchmark parameters, allocator
mode, raw outputs, and system context, making the experiment inspectable
and straightforward to rerun on different hardware. On the tested
WSL2/i7-10700K host, the THP-enabled runs reproduced the underlying
mechanism more convincingly than prior attempts: mid-run samples and
end-state summaries confirm hugepage-backed anonymous memory. The
balanced repetitions also make the result more scientifically useful
than a single fixed-order run, because they expose the run-to-run
variation that matters when the expected Redis effect is below one
percent.

The principal limitation is environmental fidelity. The original Redis
experiments were conducted on Intel Skylake Xeon servers using a
specific LLVM revision and Google's internal TCMalloc build environment.
A Docker container on a consumer desktop cannot replicate the same kernel
behavior, THP availability, physical memory fragmentation profile, NUMA
topology, or hardware counter semantics. These factors are particularly
consequential for hugepage experiments, since hugepage availability is a
system-wide property that the benchmark process does not control.

Source identity is a related concern. The public TCMalloc commit used
here is a historical approximation chosen because it still admits a clean
legacy-versus-Temeraire split; it should not be taken as evidence that
the internal artifact from the paper was reconstructed. The public
TCMalloc repository has evolved substantially beyond the OSDI~2021
codebase, and the two have diverged in ways that are difficult to
characterize from the outside.

Despite these limitations, the reproduction serves a concrete purpose.
It translates the paper's Redis methodology into explicit, runnable
steps and provides a basis for discussing which claims in systems papers
are verifiable from public artifacts and which depend on inaccessible
production context. More concretely, it makes the paper's central
technical claim testable at small scale: allocator placement decisions
can affect application performance through channels beyond the direct
cost of \texttt{malloc} and \texttt{free}.

The main empirical lesson from the repeated runs is that local Redis
results should be reported as a distribution, not as a single definitive
number. Release-off repeatedly moves in Temeraire's favor and is close to
the paper's reported scale. Release-on is more ambiguous: most runs are
near flat or modestly positive, but the targeted rerun was needed to show
that one very negative result was not reproducible. This is still useful
evidence, but it is evidence about a noisy single-host approximation,
not a confirmation of the paper's exact production-era measurements.

\section{Conclusion}

This artifact provides a narrow local reproduction path for the
Temeraire paper, deliberately scoped to the Redis case study. It does
not attempt to reproduce the fleet-scale results, which depend on
Google's internal workload diversity and infrastructure. Its purpose is
instead to make the underlying mechanism testable in a controlled,
public setting.

The experiment should be read as local evidence about one public
workload under two allocator configurations, not as a confirmation of
the paper's production claims. Under THP-enabled conditions, release-off
shows a recurring positive Temeraire signal in the same small-effect
range as the paper. Release-on is smaller and noisier; after one severe
outlier was rerun directly, the targeted result returned to a near-flat
$+0.12\%$. The final claim is therefore intentionally limited: the
artifact reproduces the Redis methodology and often lands in the paper's
effect regime, but it does not robustly reproduce the exact Redis
Table~1 statistics.

Its broader contribution is to make the mechanism concrete: allocation
affects performance not only through the speed of \texttt{malloc} and
\texttt{free}, but also through the memory layout it produces. Huge
pages can reduce TLB pressure, but they make fragmentation more
consequential. Temeraire addresses this by packing allocations so that
some hugepages remain dense while others can drain to the point where
release is worthwhile, a design goal that even a local Docker experiment
can begin to probe.

\clearpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{temeraire-seminar-report}

\end{document}
